### 2019-10-14 ch8_discussion/content.tex
update description

Despite \texttt{SCORPIUS} published on bioRxiv in 2016, discussion and usage of the pre-print has largely been limited to comparing the results of a different (linear) TI methods to those of \texttt{SCORPIUS}. Since the pre-print, many improvements have been made to improve the accuracy and robustness of predicted trajectories, to allow scaling up to datasets containing 100'000s of cells, and to providing a more user-friendly interface. Since the beginning of 2019 -- before the publication of our benchmark of TI methods -- several studies have now reported using SCORPIUS as a tool for deriving primary results\cite{damond_maphumantype_2019, cheng_multifactorialheterogeneityvirusspecific_2019, tibbitt_singlecellrnasequencing_2019, hove_singlecellatlasmouse_2019, wouters_singlecellgeneregulatory_2019}.

\section{bred: Inferring single cell regulatory networks}

%	\item We develop an algorithm for  \textbf{inferring gene regulatory interactions at a single-cell level} (Chapter~\ref{chap:bred}). We demonstrate our NI methodology by inferring regulatory interactions for 25'000 hematopoietic cells, as well as for 15'000 cancer-related bulk profiles. % TODO: was this chapter written in the end?

%\item We invent a new type of NI method capable of inferring the GRN of individual cells (Chapter~\ref{chap:bred}). We demonstrate this <yadeyade .. fill in when the chapter is actually written.>

\section{incgraph: Optimising regulatory networks}

%	\item We provide a tool for  \textbf{analysing the topological properties of large, evolving networks} and use this to iteratively optimise GRN predictions (Chapter~\ref{chap:incgraph}).

%\item Every NI method has certain topological biases. We provide a tool for analysing the topological properties of large, evolving networks and use this to iteratively optimise GRN predictions (Chapter~\ref{chap:incgraph}).

>>>>

Despite \texttt{SCORPIUS} published on bioRxiv in 2016, discussion and usage of the pre-print has largely been limited to comparing the results of a different (linear) TI methods to those of \texttt{SCORPIUS}. Since the pre-print, many improvements have been made to improve the accuracy and robustness of predicted trajectories, to allow scaling up to datasets containing 100'000s of cells, and to providing a more user-friendly interface. Since the beginning of 2019 -- before the publication of our benchmark of TI methods -- several studies have now reported using \texttt{SCORPIUS} as a tool for deriving primary results\cite{damond_maphumantype_2019, cheng_multifactorialheterogeneityvirusspecific_2019, tibbitt_singlecellrnasequencing_2019, hove_singlecellatlasmouse_2019, wouters_singlecellgeneregulatory_2019}.

\section{Inferring single cell regulatory networks}
\mycomment{TODO: fill in when chapter has been written.}
%	\item We develop an algorithm for  \textbf{inferring gene regulatory interactions at a single-cell level} (Chapter~\ref{chap:bred}). We demonstrate our NI methodology by inferring regulatory interactions for 25'000 hematopoietic cells, as well as for 15'000 cancer-related bulk profiles. % TODO: was this chapter written in the end?

%\item We invent a new type of NI method capable of inferring the GRN of individual cells (Chapter~\ref{chap:bred}). We demonstrate this <yadeyade .. fill in when the chapter is actually written.>^

\section{Optimising regulatory networks}
This project was a spin-off of Netter\cite{ruyssinck_netterrerankinggene_2016}, a tool for reranking GRN predictions using structural network properties. 
Predicted GRNs often are topologically dissimilar to real networks, and as such, Netter quantifies the topological properties of a network and makes incremental changes in order to arrive at a network with a 'better' topological profile.

The topological 'profile' of a network was defined as the frequencies of particular topological patterns called 'graphlets'. Keeping track of the frequencies of graphlets during an iterative process proved difficult at first. 
\texttt{incgraph} provides a solution to this problem (Chapter~\ref{chap:incgraph}) by making assuming that the incremental changes between two different networks are relatively small (less than 100 edge additions/removals)). Under this assumption, it is possible to calculate the differences in topological patterns significantly faster in comparison to recomputing the graphlet frequencies from scratch at every iteration.

While \texttt{incgraph} solved a very specific and interesting problem, and was eventually was published as a peer-reviewed article, it has not succeeded in making a direct impact on scientific research as of yet. Perhaps, at some point, a researcher will come across \texttt{incgraph} and find that it exactly solves his or her problem.
